---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# MicrobiotaProcess: A comprehensive R package for managing and analyzing microbiome and other ecological data within the tidy framework

If you use this work in published research, please cite:

S Xu, L Zhan, W Tang, Q Wang, Z Dai, L Zhou, T Feng, M Chen, T Wu, E Hu, G Yu. MicrobiotaProcess: A comprehensive R package for managing and analyzing microbiome and other ecological data within the tidy
framework. 04 February 2022, [PREPRINT (Version 1) available at Research Square](https://doi.org/10.21203/rs.3.rs-1284357/v1).

This repo contains source code and data to produce Supplementary Material of the above paper.

To compile the [supplementary_file_A.pdf](https://github.com/YuLab-SMU/MP_supplementary_file/blob/main/supplemental_file_A.pdf) and [supplementary_file_B.pdf](https://github.com/YuLab-SMU/MP_supplementary_file/blob/main/supplemental_file_B.pdf), please run the following command in R:

```
rmarkdown::render("supplementary_file_A.Rmd")
rmarkdown::render("supplementary_file_B.Rmd")
```

```{r echo=FALSE, results = 'hide', message = FALSE}
   library(SummarizedExperiment) # SummarizedExperiment container, a container contains one or more assays.
   library(MicrobiotaProcess) # an R tidy framework for microbiome or other related ecology data analysis.
   library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics.
   library(coin) # Conditional Inference Procedures in a Permutation Test Framework.
   library(ggnewscale) # Multiple Fill and Colour Scales in 'ggplot2'.
   library(forcats) # Helpers for reordering factor levels and tools for modifying factor levels.
   library(ggtree) # visualizing phylogenetic tree and heterogenous associated data based on grammar of graphics.
   library(ggtreeExtra) # plot associated data on the external layer based on grammar of graphics.
   library(clusterProfiler)
   library(enrichplot)
   library(MicrobiomeProfiler)
   library(igraph)
   library(aplot)
   library(tidybulk)
   library(edgeR)
   library(ggpp)
   library(broom)
   library(ggVennDiagram)
   library(shadowtext)
   library(ggrepel)
   library(patchwork)
   library(ggupset)
   library(ggnewscale)
   library(pROC)
   library(curatedMetagenomicData)
   library(randomForest)
   library(GUniFrac)
   library(matrixStats)
   library(ANCOMBC)
   library(MicrobiomeStat)
   library(metagenomeSeq)
```

Here is the output of `sessioninfo::session_info()` on the system on which [this document](https://github.com/YuLab-SMU/MP_supplementary_file/blob/main/supplemental_file_A.pdf) was compiled:

```{r echo = FALSE, warning = FALSE}
options(width = 200)
sessioninfo::session_info()
```

